Two-Dimensional Vesicle Hydrodynamics from Hydrophobic Attraction Potential

### 2021-09-22 doc/main_jfm.tex
worked on up to sec 4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figure0_Wrapper.pdf}
\end{center}
\caption{Panel (a) illustrates the domain and boundary conditions for the mobility problem. 
The color map in panels (b) - (e) is the solution $u$ of \eqref{eq:SL}. Red is for $u = 1$
and blue is for $u = 0$.
Particles self-assemble into vesicle bilayers (panel (b)) and eventually  
arrange along inner and outer leaflets. In panel (c), $C_M$ is the midplane curve,
$A$ is the enclosed area, and $L$ is the arc length.  Initially stretched vesicles relax to equilibrium 
with fluid flow across the bilayer. The study considers a single vesicle in background flows such as shear flow (panel (d))
and hydrodynamic vesicle-vesicle interactions (panel (e)).}
\label{fig:figure0}
\end{figure}

\section{Governing Equations\label{sec:governing_eqs}}
\subsection{\label{mobility}Mobility Problem}
The objective is to study the hydrodynamics of JP vesicles in background flows. 
We consider an $N_b$-many body collection of JP suspended
in a two-dimensional unbounded domain $\Omega$. 
The boundary of each particle is denoted by $\Gamma_i$ so that $\bd \Omega = \Gamma_1 \cup
\Gamma_2 \cup \cdots \cup \Gamma_{N_b}$ (Figure \ref{fig:figure0}a). 
Assuming the inertial terms are negligible, the governing equations are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $p$ is the
pressure, $\uu_\infty$ is the background flow, and $\mu$ is the constant viscosity. 
Since each particle $\Gamma_i$ with center $\aa_i$ is a rigid body, its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity. Here, $\langle x, y \rangle^{\perp} = \langle -y, x
\rangle$. Therefore, the no-slip boundary condition on each particle is
\begin{align}
\label{eq:rigid_bc}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \Gamma_i.
\end{align}
To determine the translational and angular velocities of each particle,
we define imposed forces $\FF_i$ and torques $T_i$ acting on each
particle. Since the small particles are inertialess, force balance
gives 
\begin{alignat}{2}
  \label{eq:force}
  \FF_i &- \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \mathbf{0},
  && i=1,\ldots,N_b,\\
  \label{eq:torque}
  T_i &- \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu) \, \dif s = 0, \qquad && i=1,\ldots,N_b,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor) and
$\nnu_i$ is the particle outward normal. The process of finding the
translational and angular velocities given the forces and torques is
referred to as the mobility problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Imposed Forces}
The imposed forces and torques contain two parts: hydrophobic attraction and repulsion. 
The hydrophobic attraction potential was introduced by ~\cite{Fu20} and is responsible for forming particle aggregates
that sequester their hydrophobic surface regions (Figure \ref{fig:figure0}b).
We model hydrophobic attraction by solving
the screened Laplace equation boundary value problem
\begin{alignat}{2}
  \label{eq:SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{eq:SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i,\; i=1,\ldots,N_b \\
\label{eq:SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $0 \leq f_i \leq 1$ is a material label with $f_i = 0$,
respectively $1$, representing hydrophilic, respectively hydrophobic,
portions of the surface. We assume that the $f_i$ and $\Gamma_i$ are
smooth. The parameter $\rho > 0$ is the decay length of attraction. The
forces and torques of attraction are 
\begin{align}
  \label{eq:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu \, \dif s, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu) \dif s,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)
\end{align}
is the hydrophobic stress tensor and $\gamma > 0$ is
the interfacial
tension.

@@

To calculate the hydrophobic force and torque, the gradient of the
double-layer potential~\eqref{eq:HAP_DLP} must be computed on the
boundary of each particle. The resulting integrands are singular, and
specialized quadrature would be necessary to approximate such integrals.
Alternatively, in Section~\ref{subsec:calculating_force}, we show that
the force and torque calculations can be expressed in terms of
non-singular integrals.

@@

After discretizing and applying appropriate quadrature rules, the result
is a $(2NN_b + 3N_b) \times (2NN_b + 3N_b)$ linear system that we solve
with block-diagonal preconditioned GMRES. Other BIE
formulations of the mobility problem have used single-layer potentials
(\cite{cor-gre-rac-vee2017, rac-gre2016}) or a combination of single-
and double-layer potentials (\cite{cor-vee2018}).

We validated our solver for \eqref{eq:SKIE}-\eqref{eq:mobility2} using a
single elliptical particle suspended in a background shear flow (see
\S~\ref{sec:ves_in_shear}). Hydrophobic attraction and repulsion are
zero for a single particle (see \cite{Fu20} equation 2.13). The angle of the ellipse's major axis coming
from the integral equation method was in complete agreement with the
theoretical, Jeffery orbit time-course (\cite{jef1922}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Main Theoretical Result: Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:screenedSKIE} has been solved for $\sigma$, we need to
evaluate the integrals~\eqref{eq:hydrophobicAttraction} which are the
HAP forces and torques. These integrals involve the
stress~\eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
first define
\begin{align}
  \label{eq:uidef}
  v_i(\xx) = u(\xx) - u_i(\xx),
\end{align}
where
\begin{align}
  u_i(\xx) &= \frac{1}{2\pi} \int_{\Gamma_i} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \mathbb{R}^2.
\end{align}
That is, $v_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP}
with $\Gamma_i$ excluded from $\bd\Omega$. Having defined $v_i$, we prove
\begin{theorem}
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\mathrm{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma v_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{d\nnu} \ttau.
\end{align}
\end{theorem}
The symbols $\ttau$ and $\frac{d}{ds}$ are the unit tangent and arc length
derivative for $\Gamma_i$, respectively (Figure \ref{fig:figure0}a.)  
The result 
is valid for any smooth particle shape or boundary condition. 
The advantage in using~\eqref{eq:recipforcetorque} over
using~\eqref{eq:hydrophobicAttraction} is that the components $\sigma$,
$v_i$, $\frac{d\sigma}{ds}$, and $\frac{dv_i}{ds}$ of $\JJ_i$ are smooth
functions, whereas the components of \eqref{eq:stress} are singular
integrals.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figure0_Wrapper.pdf}
\end{center}
\caption{Panel (a) illustrates the domain and boundary conditions for
  the mobility problem. The color map in panels (b)--(e) is the solution
  $u$ of \eqref{eq:SL}. Red is for $u = 1$ and blue is for $u = 0$.
  Particles self-assemble into vesicle bilayers (panel (b)) and
  eventually arrange along inner and outer leaflets. In panel (c), $C_M$
  is the midplane curve, $A$ is the enclosed area, and $L$ is the arc
  length. Initially stretched vesicles relax to equilibrium with fluid
  flow across the bilayer. The study considers a single vesicle in
  background flows such as shear flow (panel (d)) and hydrodynamic
  vesicle-vesicle interactions (panel (e)).}
\label{fig:figure0}
\end{figure}

\section{Governing Equations\label{sec:governing_eqs}}
\subsection{\label{mobility}Mobility Problem}
The objective of this work is to study the hydrodynamics of JP vesicles
in background flows. We consider an $N_b$-many body collection of JP
suspended in a two-dimensional unbounded domain $\Omega$. The boundary
of each particle is denoted by $\Gamma_i$ so that $\bd \Omega = \Gamma_1
\cup \Gamma_2 \cup \cdots \cup \Gamma_{N_b}$ (Figure
\ref{fig:figure0}a). Assuming the inertial terms are negligible, the
governing equations are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $p$ is the
pressure, $\uu_\infty$ is the background flow, and $\mu$ is the constant viscosity. 
Since each particle $\Gamma_i$ with center $\aa_i$ is a rigid body, its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity. Here, $\langle x, y \rangle^{\perp} = \langle -y, x
\rangle$. Therefore, the no-slip boundary condition on each particle is
\begin{align}
\label{eq:rigid_bc}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \Gamma_i.
\end{align}
To determine the translational and angular velocities of each particle,
we define imposed forces $\FF_i$ and torques $T_i$ acting on each
particle. Since the small particles are inertialess, force and torque
balance gives 
\begin{alignat}{2}
  \label{eq:force}
  \FF_i &- \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \mathbf{0},
  && i=1,\ldots,N_b,\\
  \label{eq:torque}
  T_i &- \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu) \, \dif s = 0, \qquad && i=1,\ldots,N_b,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor) and
$\nnu_i$ is the particle outward normal. The process of finding the
translational and angular velocities given the forces and torques is
referred to as the mobility problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Imposed Forces}
The imposed forces and torques contain two parts: hydrophobic attraction
and repulsion. The hydrophobic attraction potential was introduced
by~\cite{Fu20} and is responsible for forming particle aggregates that
sequester their hydrophobic surface regions (Figure \ref{fig:figure0}b).
We model hydrophobic attraction by solving the screened Laplace equation
boundary value problem
\begin{alignat}{2}
  \label{eq:SL}
-\rho^2 \Delta u + u &=0,            && \xx \in \Omega,\\
\label{eq:SLbc}
u(\xx) &= f_i(\xx),\qquad  && \xx \in \Gamma_i,\; i=1,\ldots,N_b, \\
\label{eq:SLff}
u &\to 0,                          &&|\xx| \to \infty,
\end{alignat}
where $0 \leq f_i \leq 1$ is a material label with $f_i = 0$,
respectively $1$, representing hydrophilic, respectively hydrophobic,
portions of the surface. We assume that both $f_i$ and $\Gamma_i$ are
smooth. The parameter $\rho > 0$ is the decay length of attraction. The
forces and torques of attraction are 
\begin{align}
  \label{eq:hydrophobicAttraction}
  \FF_i^{\text{hydro}} = \int_{\Gamma_i} {\bf T}\cdot \nnu \, \dif s, 
    \quad 
  T_i^{\text{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot ({\bf T} \cdot \nnu) \dif s,
\end{align}
where
\begin{align}
  \label{eq:stress}
\mathbf{T}
= \gamma\rho^{-1}u^2 \mathbf{I} + 2\rho\gamma \left(\tfrac{1}{2}|\nabla
  u|^2 \mathbf{I} - \nabla u  \nabla u^T\right)
\end{align}
is the hydrophobic stress tensor and $\gamma > 0$ is
the interfacial
tension.

@@

To calculate the hydrophobic force and torque, the gradient of the
double-layer potential~\eqref{eq:HAP_DLP} must be computed on the
boundary of each particle. The resulting integrands are singular, and
specialized quadrature would be necessary to approximate such integrals.
Alternatively, in Section~\ref{subsec:calculating_force}, we show how
the force and torque calculations can be expressed in terms of
non-singular integrals.

@@

After discretizing and applying appropriate quadrature rules, the result
is a $(2NN_b + 3N_b) \times (2NN_b + 3N_b)$ linear system that we solve
with block-diagonal preconditioned GMRES. Other BIE formulations of the
mobility problem use single-layer potentials (\cite{cor-gre-rac-vee2017,
rac-gre2016}) or a combination of single- and double-layer potentials
(\cite{cor-vee2018}).

We have validated our solver for~\eqref{eq:SKIE}--\eqref{eq:mobility2}
using a single elliptical particle suspended in a background shear flow
(see \S~\ref{sec:ves_in_shear}). Hydrophobic attraction and repulsion
are zero for a single particle (see \cite{Fu20} equation 2.13). The
angle of the ellipse's major axis coming from the integral equation
method agrees with the theoretical, Jeffery orbit time-course
(\cite{jef1922}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Main Theoretical Result: Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:screenedSKIE} has been solved for $\sigma$, we need to
evaluate the integrals~\eqref{eq:hydrophobicAttraction} which are the
HAP forces and torques. These integrals involve the
stress~\eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
first define
\begin{align}
  \label{eq:uidef}
  v_i(\xx) = u(\xx) - u_i(\xx),
\end{align}
where
\begin{align}
  u_i(\xx) &= \frac{1}{2\pi} \int_{\Gamma_i} \pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) \, \dif s_\yy,
    \quad \xx \in \mathbb{R}^2.
\end{align}
That is, $v_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP}
with $\Gamma_i$ excluded from $\bd\Omega$. Having defined $v_i$, we prove
\begin{theorem}
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\mathrm{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma v_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{d\nnu} \ttau.
\end{align}
\end{theorem}
The symbols $\ttau$ and $\frac{d}{ds}$ are the unit tangent and arc
length derivative for $\Gamma_i$, respectively (Figure
\ref{fig:figure0}a.) The result is valid for any smooth particle shape
or boundary condition. The advantage of
using~\eqref{eq:recipforcetorque} over
using~\eqref{eq:hydrophobicAttraction} is that the components $\sigma$,
$v_i$, $\frac{d\sigma}{ds}$, and $\frac{dv_i}{ds}$ of $\JJ_i$ are smooth
functions, whereas the components of \eqref{eq:stress} are singular
integrals.
